---
title: Traffic Data Analysis
output: html_document
---

```{r}
library(dplyr)
```

## Reading Data

```{r}
data = read.csv("all-semi-unique.csv")
dim(data)
names(data)
str(data)
summary(data)
```

Having a look on the first few rows of the data.

```{r}
head(data)
```

## Cleaning Data

### Processing Columns

Removing advertisements columns.

```{r}
data <- data %>% select(-starts_with("ad."))
ncol(data)
names(data)
```

Removing road images, colors and rode report images.

```{r}
data <- subset(data, select = -c(rd.img, rd.cl, rd.rp.rpImg, rd.rp.img))
ncol(data)
names(data)
```

Checking columns with constant features and printing their names.

```{r}
col_ct = sapply(data, function(x) length(unique(x)))
length(col_ct[col_ct == 1])
names(col_ct[col_ct == 1])
```

Removing constant-feature columns.

```{r}
data = data[, !names(data) %in% names(col_ct[col_ct == 1])]
ncol(data)
names(data)
```

### Processing Rows

Checking proportion of N/A values.

```{r}
sum(is.na(data)) / nrow(data) / ncol(data)
```

Checking number of duplicate rows.

```{r}
nrow(data) - nrow(unique(data))
```

***
TODO:  
Guess Rest of Column Meanings  
Verify Date Format  
Check for Unnecessary N/A Values  
Feature Engineering  